Ex 10: Simulation Studies

### 2022-02-15 wham_simulation/tutorial/vignette_10_wham_simulation.Rmd
fixed up some issues in the vignette

Several of the elements in the object returned by `fit_wham` are those produced by `TMB::MakeADFun`. `mod$par` are the fixed effects or parameters over which the marginal likelihood is optimized. `mod$fn` is the function that returns the (Laplace approximation of) the negative marginal log-likelihood. `mod$gr` is the function that returns the gradient and `mod$he` returns the hessian. `mod$report` provides a list of output defined by REPORT() calls in the compiled cpp code and is already evaluated and provided by `mod$rep`. When available, `mod$sdrep` provides the hessian based standard errors for estimates from the model which is generated by `TMB::sdreport`.

@@

```
Note that the variance of the M random effects we are assuming is the same as the default value for estimation. Note also, that the mean of the M autoregressing random effects is defined by the  `initial_means` component of the `M` argument to `prepare_wham_input` and it is not estimated by default. To estimate the mean of the M random effects at age, or just estimate a constant M without random effects, specify the `M$est_ages` argument to `prepare_wham_input`. Setting the correlation parameters when `M$re = "ar_a"` or  `M$re = "ar_y"` is analogous to the description for numbers at age random effects.

@@

```{r, eval=FALSE}
stock_om_info = make_basic_info()

@@

When environmental effects are considered, there are assumptions to be made about the variance and any autocorrelation parameters for the random effects for the latent environmental process and the variance of the observation errors. The latter are specified in the the `ecov$log_sigma` argument to `prepare_wham_input`, but those for the random effects (`Ecov_process_pars`) must be specified after the call to `prepare_wham_input` like the other random effects described above. The value for the effect of the covariate on the population (`Ecov_beta`) must also be specified this way. The columns of `Ecov_process_pars` correspond to each environmental covariate being modeled and the number of rows is 3. If `ecov$process_model = "rw"` only the first two elements are used and represent the value in the first year and the conditional standard devation parameter of the random walk. When `ecov$process_model = "ar1"` the three elements correspond to mean, variance, and correlation parameters.

@@

As long as there is only a single environmental effect, we can change the appropriate element of the `Ecov_beta` array by using the `map$Ecov_beta` specification returned by `prepare_wham_input`. The `map$Ecov_beta` will be all NAs except for the configured effects on the population and when there is just one there is no ambiguity. Otherwise, the 4 correct indices of the `Ecov_beta` array must be referenced appropriately for each effect.

@@

#
NAA_re_om = list(
  N1_pars = exp(10)*exp(-(0:(length(stock_om_info$ages)-1))*M_om$initial_means[1]),
  sigma = "rec", #random about mean
  recruit_model = 2, #random effects with a constant mean
  recruit_pars = exp(10)
)

ecov_om = list(
    label = "Climate variable 1",
    process_model = "rw",
    mean = matrix(0, length(stock_om_info$years), 1), #observations which will be simulated later
    logsigma = log(0.2), 
    lag = 0,
    years = stock_om_info$years, 
    use_obs = matrix(1,length(stock_om_info$years),1),  
    where = "recruit",
    how = 1)

@@

#par(mfrow = c(2,1))
#plot(stock_om$years, sim_Ecov$Ecov_x, type = 'l', ylab = "Covariate", xlab = "Year")
#plot(stock_om$years, sim_Ecov$NAA[,1], type = 'l', ylab = "Recruitment", xlab = "Year")

@@

```{r, eval=TRUE}
M_em = list(initial_means = rep(0.3, length(stock_om_info$ages)))

>>>>

Several of the elements in the object returned by `fit_wham` are those produced by `TMB::MakeADFun`. `mod$par` are the fixed effects or parameters over which the marginal likelihood is optimized. `mod$fn` is the function that returns the (Laplace approximation of) the negative marginal log-likelihood. `mod$gr` is the function that returns the gradient and `mod$he` returns the hessian. `mod$report` is a function that provides a list of output defined by REPORT() calls in the compiled cpp code and is already evaluated and provided by `mod$rep`. When available, `mod$sdrep` provides the hessian based standard errors for estimates from the model which is generated by the`TMB::sdreport` function.

@@

```
Note that the variance of the M random effects we are assuming is the same as the default value for estimation. Note also, that the mean of the M autoregressing random effects is defined by the  `initial_means` component of the `M` argument to `prepare_wham_input` and it is not estimated by default. To estimate the mean of the M random effects at age, or just estimate a constant M without random effects, specify the `M$est_ages` argument to `prepare_wham_input`. There is a difference from numbers at age in how random effects for M are specified. If `M$re = "ar_a"` then there are only `n_ages` random effects that are constant over time. Similarly for `M$re = "ar_y"`, there are `n_years_model` random effects that are used for each age class. However, Setting the correlation parameters when `M$re = "ar_a"` or  `M$re = "ar_y"` is analogous to the description for numbers at age random effects. To get `ar_a` or `ar_y` models analogous to numbers at age, one would have to specify `M$re = "2dar1"` and change `map$M_repars` to fix the appropriate parameter fixed at 0. To force random effects for certain age classes to be the same, one would also have to configure `map$M_re` appropriately.

@@

```{r, eval=TRUE}
stock_om_info = make_basic_info()

@@

When environmental effects are considered, there are assumptions to be made about the variance and any autocorrelation parameters for the random effects for the latent environmental process and the variance of the observation errors. The latter are specified in the the `ecov$log_sigma` argument to `prepare_wham_input`, but those for the random effects (`Ecov_process_pars`) must be specified after the call to `prepare_wham_input` like the other random effects described above. The value for the effect of the covariate on the population (`Ecov_beta`) must also be specified this way. The columns of `Ecov_process_pars` correspond to each environmental covariate being modeled and the number of rows is 3. If `ecov$process_model = "rw"` only the first two elements are used and represent the value in the first year and the conditional standard deviation parameter of the random walk. When `ecov$process_model = "ar1"` the three elements correspond to mean, variance, and correlation parameters.

@@

As long as there is only a single environmental effect, we can change the appropriate element of the `Ecov_beta` array by using the `map$Ecov_beta` specification returned by `prepare_wham_input`. The `map$Ecov_beta` will be all `NA` except for the configured effects on the population and when there is just one there is no ambiguity. Otherwise, the 4 correct indices of the `Ecov_beta` array must be referenced appropriately for each effect.

@@

NAA_re_om = list(
  N1_pars = exp(10)*exp(-(0:(length(stock_om_info$ages)-1))*M_om$initial_means[1]),
  sigma = "rec", #random about mean
  recruit_model = 2, #random effects with a constant mean
  recruit_pars = exp(10)
)

ecov_om = list(
    label = "Climate variable 1",
    process_model = "ar1",
    mean = matrix(0, length(stock_om_info$years), 1), #observations which will be simulated later
    logsigma = log(0.2), 
    lag = 0,
    years = stock_om_info$years, 
    use_obs = matrix(1,length(stock_om_info$years),1),  
    where = "recruit",
    how = 1)

@@

plot(stock_om$years, sim_Ecov$Ecov_x, type = 'l', ylab = "Covariate", xlab = "Year")
plot(stock_om$years, sim_Ecov$NAA[,1], type = 'l', ylab = "Recruitment", xlab = "Year")

@@

M_em = list(
  model = "constant", 
  initial_means = 0.3)
